### 2022-04-24 outputs/paper/paper.Rmd
reference and introduction

---
title: "Religion and Its Influence on Social Issues^[Code and data are available at: https://github.com/dudqo/religion_survey]"
subtitle: "Possible factors effecting 2016 election"
author: 
  - Young Suk
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "This paper has been examined possible variables that influenced U.S citizens' choice for the 2016 U.S presidential election. We have found that there exists significantly different opinions regarding gender, wealth, race and age. This is important in a sesne that it shows thought trends as well as crucial decision indicators for voting in U.S.,With further survey at the end of this paper, we are going to explore whether this is happening only in U.S or not."
output:
  bookdown::pdf_document2:
    fig_caption: yes
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(dplyr)
library(knitr)
```

# Introduction
Religion has been major part of human history around the world, and for many countries, religion is something they heavily rely on even in modern days.For example, European countries such as Spain and Italy which heavily relies on catholic, and there are many middle eastern countries that heavily rely on Islam and Hinduism. While it is reported that the number of religious people is declining quickly, religion still has big social influence in many aspects of the society in the United States, and this paper aims to research the social influence of popular religions in the United States and how the opinions of different religions differ in the context of social issues and government policies regarding them.

This research will focus on some social issues in the United States and how opinions of people of different religion differs.

The dataset was obtained from U.S GSS survey, by only selecting the survey questions that are related to the topic of this research. The data used includes personal profile data of survey respondents such as age, gender, religion, race, income and data on questions regarding their opinion on popular social issues in the United States such as abortion, gay marriage and government policies regarding welfare, racism, and etc. This research found that despite the people's different views on different religion and the differences in the religions themselves, when it comes to the popular social issues, their opinions are mostly inline with each other. Some different views do exists and which will be discussed in the later part of the paper.

This report was created using `R` [@citeR] and its libraries `tidyverse` [@tidyverse], `dplyr` [@dplyr], were used to process and clean the data, and `ggplot` [@ggplot] `bookdown` [@bookdown] and `knitr` [@knitr] was used to format this report.

# Data
```{r, include=FALSE, echo=FALSE}
# Read the reduced and cleaned data, modification done in data-preparation.r file
reduced_data <- read_csv("outputs/data/prepared_gss.csv")
# Change the order of variables for graphing purposes
reduced_data$Family_income <- factor(reduced_data$Family_income, 
                                     levels = c("Under $10,000","$10,000 ~ $19,999","$20,000 ~ $39,999","$40,000 ~ $59,999", 
                                                "$60,000 ~ $89,999","$90,000 ~ $129,999","$130,000 ~ $169,999","$170,000 or over"))
reduced_data$Immigration <- factor(reduced_data$Immigration, levels = c("Strongly disagree","Disagree", "Neither","Agree","Strongly agree"))
```

The Original data is from survey by US GSS, which contains data of public opinions of USA citizens on various subjects. The data is from 2021 survey which was conducted online due to global pandemic by COVID19. For the purpose of this research, the data related to personal profiles, survey respondents' opinion on government policies and religion has been extracted from the raw survey data.

The religion data has very detailed options that includes religions have very low For the simplicity of this research, protestants and Christians were grouped together under Christians and all the non-popular religion has been grouped under "other".

Looking at the figures, some expected trends from the data can be observed:

* As age gets younger, the percentage of people who does not have any religion increases. With steady increase in nonreligious as age group gets younger, people in the age group of 80-89 have very few people who are not religious. If we look at 20-29 years old, we can see that about half of them does not have any religion.
* There are far more white population who participated in the survey than any other race. This is as expected since white population in the United States far outnumbers any other race.
* Christian is most popular religion in the United States.

```{r fig1,echo=FALSE, fig.cap = "Distribution of survey respondents' age group."}
reduced_data %>%
  drop_na(age_group) %>%
  ggplot(aes(x = age_group, fill = relig)) +
  geom_bar() +
  coord_flip() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Survey Respondents' Age") +
  theme(legend.title = element_blank())
```

```{r fig2, fig.cap= "Distribution of survey respondents' race.", echo = FALSE}
reduced_data %>%
  ggplot(aes(x = race, fill = relig)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Survey Respondents' Race",
       ) +
  theme(legend.position = "none")
```

```{r fig3, echo = FALSE, fig.cap= "Distribution of survey respondents' gender"}
reduced_data %>%
  ggplot(aes(x = gender, fill = relig)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Gender and Opinion on Importance of Being Born as Male or Female",
       ) +
  theme(legend.title = element_blank()) +
  coord_flip() +
  facet_wrap(~opsex)
```
```{r fig4, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
reduced_data %>%
  ggplot(aes(x = relig, fill = born)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Number of Survey Respondents' Born in U.S",
       ) +
  coord_flip() +
  theme(legend.title = element_blank())
```

```{r fig5, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
reduced_data %>%
  ggplot(aes(x = relig, fill = Family_income)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Survey Respondents' Religion",
       ) +
  coord_flip() +
  theme(legend.title = element_blank())
```
```{r fig6, echo=FALSE, fig.cap = "Distribution of religion of survey respondents by family income"}
reduced_data %>%
  ggplot(aes(x = relig, fill = relig)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Survey Respondents' Family Income by Religion",
       ) +
  coord_flip() +
  theme(legend.title = element_blank())
```

```{r fig7, echo=FALSE, fig.cap = "Survey respondents' opinion on governmemt policy on improving coditions of black population"}
reduced_data %>%
  drop_na(nation_race) %>%
  ggplot(aes(x = relig, fill = nation_race)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Is the Government spending enough on improving coditions of Black People?",
       ) +
  coord_flip() +
  theme(legend.title = element_blank())
```
```{r fig8, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
reduced_data %>%
  drop_na(nation_welfare) %>%
  ggplot(aes(x = relig, fill = nation_welfare)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Is the Government Spending Enough on Welfare?",
       ) +
  coord_flip() +
  theme(legend.title = element_blank())
```

```{r fig9, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
reduced_data %>%
  drop_na(Assistance_to_the_poor) %>%
  ggplot(aes(x = relig, fill = Assistance_to_the_poor)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Is the Government Spending Enough to Asssist the Poor?",
       ) +
  coord_flip() +
  theme(legend.title = element_blank())
```
```{r fig10, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
reduced_data %>%
  drop_na(social_security) %>%
  ggplot(aes(x = relig, fill = social_security)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Is the Government Spending Enough on Social Security?",
       ) +
  coord_flip() +
  theme(legend.title = element_blank())
```
```{r fig11, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
reduced_data %>%
  drop_na(natcrime) %>%
  ggplot(aes(x = relig, fill = natcrime)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Is the Government Spending Enough to Halt Increasing Crime Rate?",
       ) +
  coord_flip() +
  theme(legend.title = element_blank())
```
```{r fig12, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
reduced_data %>%
  drop_na(Immigration) %>%
  ggplot(aes(x = relig, fill = Immigration)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Should the Government Limit Immigration to Protect U.S Culture?",
       ) +
  coord_flip() +
  theme(legend.title = element_blank())
```

```{r fig13, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
reduced_data %>%
  drop_na(religinf) %>%
  ggplot(aes(x = relig, fill = religinf)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Would the U.S be a better country if religion had less influence?",
       ) +
  coord_flip() +
  theme(legend.title = element_blank())
```
```{r fig14, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
reduced_data %>%
  drop_na(homo_marriage) %>%
  ggplot(aes(x = relig, fill = homo_marriage)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Do You Agree With Same Sex Marriage?",
       ) +
  coord_flip() +
  theme(legend.title = element_blank())
```
```{r fig15, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
reduced_data %>%
  drop_na(abortion) %>%
  ggplot(aes(x = relig, fill = abortion)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Do you agree with abortion for any reason?",
       ) +
  coord_flip() +
  theme(legend.title = element_blank())
```

\newpage

# Results
To study more on gender and the opinions on importance of being born as certain gender, linear model was created to study this matter.

As for the opinions on abortion, we can see an interesting result where only Christians and Catholics show significant amount of negativity towards abortion compared to other religions.

# Discussion

## Summary
In american society, general public's opinions on different religion varies significantly. While religions such as Christianity and Catholicism has been around since the beginning of history of the United States as a country, there are some religions such as Islam and Buddhism that Americans still are not very familiar with. With the unfamiliarity and some negative news involving a religion, the views on a specific religion be heavily influenced by stereotypes. We can observe from the u.S GSS survey that the opinions between people of different religion are not that different.

## Different Opinions between different religions.

## Weaknesses and how we should proceed in the future
As the survey was conducted in the United States, the data lacks diversity of religions and most of survey respondents' are Christians or Catholics. For more accurate and in depth research, more data from people of other religions are required. Also, the opinions regarding social issues could be influenced by the location they live. So data from different countries where Christianity or Catholicism is not the main religion could be very helpful for this research.

\newpage

# Appendix

\newpage

# References